---
title: "Getting Started: An SPPH 604 Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started: An SPPH 604 Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
link-citations: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette walks through a complete complex-survey analysis with `svyTable1`,
in the order you would tackle a [SPPH 604 / EpiMethods](https://ehsanx.github.io/EpiMethods/)
final project: build the design, describe the sample, fit and check a model,
probe an interaction, and look at survival. Every step uses the bundled
`nhanes_mortality` dataset (a processed NHANES cohort of diabetic adults), so the
whole vignette runs with no downloads.

```{r packages, message = FALSE, warning = FALSE}
library(svyTable1)
library(survey)
library(knitr)

options(survey.lonely.psu = "adjust")  # required for correct NHANES-style SEs
```

## 1. Build the survey design

The single most consequential mistake in survey analysis is to filter the data
to a subgroup *before* building the design -- that throws away the information
needed for correct subgroup standard errors. `svydesign_build()` validates your
inputs and, when you want a subpopulation, does it the right way (by subsetting
the design).

```{r design}
data(nhanes_mortality)
nhanes_mortality$htn01 <- as.numeric(nhanes_mortality$htn == "Yes")

design <- svydesign_build(
  data = nhanes_mortality,
  ids = "psu", strata = "strata", weights = "survey_weight"
)
```

## 2. Describe the sample (Table 1)

`svytable1()` builds a survey-weighted "Table 1". In the default `"mixed"` mode
each categorical cell shows the **unweighted count** (the real sample size, which
governs precision [@seidenberg2023preferred]) together with the **weighted
percentage** (the population estimate).

```{r table1}
tab1 <- svytable1(
  design = design,
  strata_var = "htn",
  table_vars = c("age", "sex", "race", "smoking", "bmi.cat")
)
knitr::kable(tab1, caption = "Table 1, stratified by hypertension status")
```

> **How to read this.** A weighted percentage describes the *population*; the
> unweighted *n* tells you how much data is behind it. With `reliability_checks
> = TRUE`, cells that fail the NCHS reliability rules are replaced by `*`
> [@nhanes_reliability_estimates].

This table drops straight into a report with `knitr::kable()` (HTML or, when you
knit to Word, a native Word table) -- no extra packages required.

## 3. Fit a model and check it

```{r model}
fit <- svyglm(htn01 ~ age + sex + smoking,
              design = design, family = quasibinomial())
```

**Coefficient reliability** with `svydiag()`:

```{r diag}
knitr::kable(svydiag(fit), digits = 3,
             caption = "Coefficient reliability diagnostics")
```

**Goodness of fit** with `svygof()` (Archer-Lemeshow; @archer2006goodness):

```{r gof}
knitr::kable(svygof(fit, design),
             caption = "Archer-Lemeshow goodness-of-fit test")
```

> **How to read this.** A *non-significant* goodness-of-fit p-value is consistent
> with adequate fit -- it does **not** prove the model is correctly specified.

**Discrimination** with `svyAUC()` (needs a replicate-weights design; the CI is
built on the logit scale so it stays in `[0, 1]`):

```{r auc}
rep_design <- as.svrepdesign(design)
fit_rep <- svyglm(htn01 ~ age + sex + smoking,
                  design = rep_design, family = quasibinomial())
knitr::kable(svyAUC(fit_rep, rep_design), digits = 3,
             caption = "Design-correct AUC")
```

> **How to read this.** AUC measures *discrimination* (ranking), not calibration
> or causal validity. An AUC near 0.5 means little discriminatory value.

## 4. Probe an interaction

Does the association between insulin use and hypertension differ by sex? Fit an
interaction model and summarise the additive interaction (RERI, AP, S):

```{r interaction}
int_fit <- svyglm(htn01 ~ sex * insulin + age,
                  design = design, family = quasibinomial())

add_int <- addintlist(int_fit, factor1_name = "sex", factor2_name = "insulin",
                      measures = "all", ci_method = "mover")
knitr::kable(add_int, digits = 3,
             caption = "Additive interaction (MOVER confidence intervals)")
```

> **How to read this.** RERI/AP/S are risk-ratio-scale measures. The `Scale`
> column shows they were computed here from **odds ratios**, which approximate
> risk ratios only when the outcome is rare; interpret with care for common
> outcomes [@knol2012recommendations]. We used `ci_method = "mover"` for the
> better-coverage RERI interval (@archer2006goodness gives background on survey
> diagnostics; the MOVER method is from Zou, 2008).

`reportint()` assembles joint, stratum-specific, additive, and multiplicative
panels into a single report; `plotint()` visualises interactions involving a
continuous moderator. See `vignette("v3-interaction-analysis")`.

## 5. Survival analysis

Survey-weighted Kaplan-Meier curves with a number-at-risk table
(`se = FALSE` skips the slower confidence bands):

```{r km, fig.width = 7, fig.height = 5}
library(survival)

km <- svykmplot(
  Surv(stime, status) ~ caff,
  design = subset(design, sex == "Female"),
  time_unit = "days",
  time_breaks = seq(0, 240, by = 60),
  show_pval = TRUE,
  se = FALSE
)
km$plot
```

The constant-effect (proportional-hazards) assumption for a Cox model can be
checked visually with `svycoxph_CE()`; the multiply-imputed version is
`svycoxph_CE_mi()`. See `vignette("v4-survival-analysis")`.

## Where to go next

| Task | Function(s) | Vignette |
|------|-------------|----------|
| Descriptive tables | `svytable1()` | `v1-svyTable1-basics` |
| Model diagnostics | `svydiag()`, `svygof()`, `svyAUC()` | `v2-model-diagnostics` |
| Interaction analysis | `addintlist()`, `reportint()`, `plotint()` | `v3-interaction-analysis` |
| Survival analysis | `svykmplot()`, `svycoxph_CE()` | `v4-survival-analysis` |
| Multiple imputation | `svypooled()` | `v5-multiple-imputation` |

See `?svyTable1` for the full function index, and the
[EpiMethods book](https://ehsanx.github.io/EpiMethods/) for the underlying
methods.

## References
