---
bibliography: references.bib
link-citations: true
output: "rmarkdown::html_vignette"
title: Creating Descriptive Table 1 Summaries with svyTable1
vignette: >
  %\VignetteIndexEntry{Creating Descriptive Table 1 Summaries with svyTable1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `svyTable1` package provides a suite of functions for creating
publication-ready tables from complex analytical results. This vignette
focuses on its core capability: generating descriptive "Table 1"
summaries using the `svytable1()` function.

## Describing Complex Survey Data

The `svytable1()` function defaults to `mode = "mixed"`. This approach
is considered a best practice in high-quality survey research because it
promotes transparency.

When you present a table in mixed mode, two types of information are
shown for categorical variables:

-   **Unweighted N:** The actual number of survey respondents in that
    category. This indicates precision [@seidenberg2023preferred].
-   **Weighted Percentage:** The proportion of the target population
    estimated to fall in that category, accounting for survey design
    [@national1994plan].

This structure prevents common reporting problems such as showing weighted counts (which do not represent real sample sizes) or hiding sparse categories that contribute instability.

For continuous variables, the function reports weighted means and standard deviations. It also offers optional reliability checks derived from National Center for Health Statistics (NCHS) guidance.

## Example: The Bundled `nhanes_mortality` Cohort

The examples below use the bundled `nhanes_mortality` dataset (a processed NHANES cohort of diabetic adults). The code illustrates both incomplete-data situations and complete-case summaries, stratified by hypertension status (`htn`).

### 1. Data Preparation

```{r packages, message = FALSE, warning = FALSE}
library(svyTable1)
library(survey)
library(knitr)

options(survey.lonely.psu = "adjust")
```


```{r data}
data(nhanes_mortality)

# This dataset is complete; induce some missingness to illustrate how
# svytable1() reports it.
set.seed(42)
nhanes_miss <- nhanes_mortality
nhanes_miss$smoking[sample(nrow(nhanes_miss), 150)] <- NA
nhanes_miss$bmi.cat[sample(nrow(nhanes_miss), 100)] <- NA

design_miss <- svydesign(
  id = ~psu,
  strata = ~strata,
  weights = ~survey_weight,
  nest = TRUE,
  data = nhanes_miss
)
```

### 2. Generating Tables

#### Example A: Handling Variables with Missing Data

Variables in large health surveys often contain missing values. `svytable1()` automatically identifies missingness and calculates unweighted Ns that reflect real denominators. Weighted percentages are based only on cases with observed data, consistent with standard survey practice.

```{r miss}
vars_with_missing <- c("age", "sex", "race", "smoking", "bmi.cat")

table_with_missing <- svytable1(
  design = design_miss,
  strata_var = "htn",
  table_vars = vars_with_missing
)

knitr::kable(
  table_with_missing,
  caption = "Table 1: Summarizing Variables with Missing Data"
)
```

#### Example B: Summarizing Complete Data

If you prefer to restrict the analysis to respondents with complete data, you can supply a design built on the complete dataset. The function then calculates summary statistics without any missing categories.

```{r complete}
design_complete <- svydesign(
  id = ~psu,
  strata = ~strata,
  weights = ~survey_weight,
  nest = TRUE,
  data = nhanes_mortality
)

table_without_missing <- svytable1(
  design = design_complete,
  strata_var = "htn",
  table_vars = c("age", "sex", "race")
)

knitr::kable(
  table_without_missing,
  caption = "Table 2: Summarizing Variables with No Missing Data"
)
```

### 3. Checking Estimate Reliability (NCHS Standards)

Federal statistical agencies discourage reporting proportions that are unstable or based on very small samples. When you request `reliability_checks = TRUE`, `svytable1()` returns:

- The formatted table with unreliable estimates suppressed (shown as `*`).

- A companion metrics table showing which estimates failed reliability standards.

These checks follow NCHS guidance for the evaluation of proportions in complex surveys [@nchs_reliability_estimates].

```{r rel}
results_list <- svytable1(
  design = design_miss,
  strata_var = "htn",
  table_vars = vars_with_missing,
  reliability_checks = TRUE,
  return_metrics = TRUE
)

knitr::kable(
  results_list$formatted_table,
  caption = "Table 3: Table with NCHS Reliability Checks Applied"
)
```

The companion metrics table records, for every evaluated cell, the unweighted
*n*, design degrees of freedom, design effect, effective sample size,
confidence-interval bounds, and the specific reliability rules that were
triggered. Filtering it to the suppressed cells explains each `*`:

```{r relmetrics}
suppressed <- results_list$reliability_metrics[
  results_list$reliability_metrics$suppressed == TRUE, ]

knitr::kable(
  head(suppressed, 10),
  caption = "Reliability Metrics for Suppressed Estimates",
  digits = 2,
  row.names = FALSE
)
```

For numeric variables, the relevant rule is the relative standard error
(RSE = standard error / estimate); an estimate is flagged when its RSE reaches
30 percent, following NCHS guidance.

## References
