---
bibliography: references.bib
link-citations: true
output: "rmarkdown::html_vignette"
title: Diagnosing Survey Regression Models with svyTable1
vignette: >
  %\VignetteIndexEntry{Diagnosing Survey Regression Models with svyTable1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette covers functions in the `svyTable1` package for diagnosing
fitted survey regression models:

-   `svydiag()`: Checks reliability of regression coefficients.\
-   `svygof()`: Performs Archer-Lemeshow goodness-of-fit testing for
    survey logistic models [@archer2006goodness].\
-   `svyAUC()`: Computes design-correct AUC for predictive performance.

These tools provide a structured way to assess whether model coefficients are precise, whether the logistic model fits observed data adequately, and whether predicted probabilities discriminate well between groups.

## 1. Data and Model Preparation

The examples use the bundled `nhanes_mortality` dataset (a processed NHANES
cohort of diabetic adults). A survey-weighted logistic regression model is
fitted to predict hypertension (`htn`). This model is used throughout the
vignette to illustrate each diagnostic function.

```{r packages, warning = FALSE, message = FALSE}
library(svyTable1)
library(survey)
library(knitr)
library(ggplot2)
```


```{r data}
data(nhanes_mortality)
nhanes_mortality$htn01 <- as.numeric(nhanes_mortality$htn == "Yes")

adult_design <- svydesign(
  id = ~psu,
  strata = ~strata,
  weights = ~survey_weight,
  nest = TRUE,
  data = nhanes_mortality
)

fit_htn <- svyglm(
  htn01 ~ age + sex + smoking,
  design = adult_design,
  family = quasibinomial()
)
```

## 2. Regression Coefficient Diagnostics (`svydiag()`)

`svydiag()` produces diagnostics for regression coefficients, including
p-values, precision measures, and relative standard errors.

```{r diag}
diagnostics_table <- svydiag(fit_htn)

knitr::kable(
  diagnostics_table,
  caption = "Reliability Diagnostics for the Hypertension Model",
  digits = 3
)
```

### Interpreting the `svydiag()` Table

-   **Estimate**: Coefficient on the log-odds scale.\
-   **p.value**, **is_significant**: Indicate statistical significance.\
-   **CI_Width**: Reflects the precision of the estimate.\
-   **RSE_percent**, **is_rse_high**: Highlight limitations of RSE when
    estimates are near zero.

## 3. Goodness-of-Fit Test (`svygof()`)

`svygof()` implements the Archer and Lemeshow goodness of fit procedure for complex survey logistic regression [@archer2006goodness]. This test adapts the Hosmer and Lemeshow approach to designs with weights, strata, and clustering.

The function groups respondents by predicted risk and compares observed and expected outcomes while incorporating survey weights. A small p-value provides evidence that predicted risks diverge from observed risks in one or more groups. Note that `svygof()` expects a standard `svydesign` object (not a replicate-weights design).

This test helps identify lack of fit that may result from omitted interactions or nonlinearities in the model.

```{r gof}
gof_results <- svygof(fit_htn, adult_design)

knitr::kable(
  gof_results,
  caption = "Archer-Lemeshow Goodness-of-Fit Test for the Hypertension Model"
)
```

A non-significant p-value is consistent with adequate fit across predicted
risk groups; it does not prove the model is correctly specified.

## 4. Model Performance (`svyAUC()`)

`svyAUC()` estimates the area under the ROC curve using replicate-weight designs. This produces an AUC estimate and confidence interval that appropriately reflect survey sampling. The confidence interval is built on the logit scale, so it always stays within `[0, 1]`.

The function also returns data for constructing ROC curves. A plot of the ROC curve, along with the AUC estimate, demonstrates how well the model discriminates between respondents with and without hypertension.

A value close to 0.5 indicates little discriminatory value, while higher values indicate stronger predictive performance.

```{r auc}
rep_design <- as.svrepdesign(adult_design)

fit_htn_rep <- svyglm(
  htn01 ~ age + sex + smoking,
  design = rep_design,
  family = quasibinomial()
)

auc_results_list <- svyAUC(fit_htn_rep, rep_design, plot = TRUE)

knitr::kable(
  auc_results_list$summary,
  caption = "Design-Correct AUC for the Hypertension Model"
)

ggplot(auc_results_list$roc_data, aes(x = FPR, y = TPR)) +
  geom_line(linewidth = 1) +
  geom_abline(linetype = "dashed") +
  labs(
    title = "Survey-Weighted ROC Curve",
    subtitle = paste0("AUC = ", round(auc_results_list$summary$AUC, 3)),
    x = "1 - Specificity (FPR)",
    y = "Sensitivity (TPR)"
  ) +
  theme_minimal()
```

## References
