---
title: "Analyzing Interaction Effects with svyTable1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing Interaction Effects with svyTable1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
link-citations: yes
---

```{r setup}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to analyze and present interaction effects using the `svyTable1` package. This includes:

- Calculating measures of additive interaction (`addintlist()`)
- Estimating joint and simple effects (`jointeffects()`, `inteffects()`)
- Generating automated multi-panel reports (`reportint()`)
- Visualizing interactions involving continuous variables (`plotint()`)

We use the bundled `nhanes_mortality` dataset (a processed NHANES cohort of diabetic adults) throughout.

## Analyzing Interaction Effects (Categorical)

The `svyTable1` package includes several functions to help analyze and present interaction effects between two categorical variables in survey-weighted regression models (`svyglm` or `svycoxph`). These functions facilitate demonstrating the equivalence of different modeling approaches and calculating measures of additive interaction.

### Example: Interaction between Sex and Insulin Use on Hypertension

For two categorical variables, `svyTable1` provides functions that summarize joint, simple, and additive interaction effects from survey-weighted logistic or Cox models. These summaries help demonstrate the connection between the saturated interaction model and the equivalent joint-indicator model.

## 1. Data Preparation and Model Fitting

```{r packages, warning = FALSE, message = FALSE}
library(svyTable1)
library(survey)
library(knitr)
library(emmeans)
library(ggplot2)

options(survey.lonely.psu = "adjust")
```

The example examines whether the association between insulin use and hypertension differs by sex. A survey-weighted logistic model with an interaction term is fitted, and a matched joint-indicator model is created to demonstrate equivalence.

```{r data}
data(nhanes_mortality)
nhanes_mortality$htn01 <- as.numeric(nhanes_mortality$htn == "Yes")

design <- svydesign(
  id = ~psu,
  strata = ~strata,
  weights = ~survey_weight,
  nest = TRUE,
  data = nhanes_mortality
)

# Saturated interaction model
interaction_model_logit <- svyglm(
  htn01 ~ sex * insulin + age,
  design = design,
  family = quasibinomial()
)

# Equivalent joint-indicator model
nhanes_mortality$sex_insulin <- interaction(
  nhanes_mortality$sex, nhanes_mortality$insulin, sep = "_"
)
design_joint <- svydesign(
  id = ~psu, strata = ~strata, weights = ~survey_weight,
  nest = TRUE, data = nhanes_mortality
)
joint_model_logit <- svyglm(
  htn01 ~ sex_insulin + age,
  design = design_joint,
  family = quasibinomial()
)

f1_levels <- c("Male", "Female")  # reference level first
f2_levels <- c("No", "Yes")       # reference level first
```

## 2. Additive Interaction Measures

`addintlist()` computes standard measures of additive interaction. These include:

- RERI (relative excess risk due to interaction)

- AP (attributable proportion due to interaction)

- S (synergy index)

These measures quantify departure from additivity. Note that for logistic models they are computed from odds ratios, which approximate risk ratios only when the outcome is rare; interpret them with care for common outcomes.

```{r add}
all_interactions_table <- addintlist(
  model = interaction_model_logit,
  factor1_name = "sex",
  factor2_name = "insulin",
  measures = "all",
  digits = 3
)

knitr::kable(all_interactions_table,
  caption = "Additive Interaction Measures (RERI, AP, S) for Sex x Insulin on Hypertension",
  digits = 3)
```

## 3. Joint Effects

`jointeffects()` estimates the combined effect for each combination of the two categorical variables. The function reports effects on either the ratio scale or the log scale and is useful for reporting the full pattern of interaction across subgroups.

```{r joint}
joint_effects_ratio <- jointeffects(
  interaction_model = interaction_model_logit,
  factor1_name = "sex",
  factor2_name = "insulin",
  scale = "ratio",
  digits = 2)

knitr::kable(joint_effects_ratio,
  caption = "Joint Effects (OR Scale) Estimated from Interaction Model",
  digits = 2)
```

## 4. Simple Effects

`inteffects()` provides stratum-specific effects: the effect of insulin within each sex, and the effect of sex within each insulin category. This helps identify which group-specific comparisons drive the interaction.

```{r simpeff}
simple_effects_ratio <- inteffects(
  joint_model = joint_model_logit,
  joint_var_name = "sex_insulin",
  factor1_name = "sex",
  factor2_name = "insulin",
  factor1_levels = f1_levels,
  factor2_levels = f2_levels,
  level_separator = "_",
  scale = "ratio",
  digits = 2)

knitr::kable(simple_effects_ratio,
  caption = "Simple Effects (OR Scale) Estimated from Joint Model",
  digits = 2)
```

## 5. Automated Multi-Panel Report

`reportint()` compiles the results into a structured multi-panel summary that includes joint effects, simple effects, additive interaction measures, multiplicative interaction estimates, and model details.

```{r auto}
interaction_report <- reportint(
  interaction_model = interaction_model_logit,
  factor1_name = "sex",
  factor2_name = "insulin",
  output = "list"
)

knitr::kable(interaction_report$joint_effects,
  caption = "Panel A (Joint Effects) from reportint()")

knitr::kable(interaction_report$additive_interaction,
  caption = "Panel C (Additive Interaction) from reportint()")

knitr::kable(interaction_report$Interaction_report,
  caption = "Publication Report: Full Interaction Analysis",
  align = 'l')
```

This panel format follows reporting recommendations [@knol2012recommendations].

## Visualizing Interaction Effects with `plotint`

`plotint()` produces plots of predicted outcomes across values of a continuous predictor for selected levels of a continuous or categorical moderator. The function uses survey-weighted marginal estimates, supports both model- and data-driven moderator scales, and returns the underlying plot data and a plot object.

This approach is helpful when evaluating interactions between continuous variables, where reporting all numerical contrasts is impractical. Here we model total caloric intake as a function of an age-by-carbohydrate interaction.

```{r viz}
model_cal_interaction <- svyglm(
  cal.total ~ age * carbohyd + sex,
  design = design,
  family = gaussian()
)

plot_data_list <- plotint(
  model = model_cal_interaction,
  effect = "age",
  moderator = "carbohyd",
  data = nhanes_mortality,
  mod_scale = "quantile",
  type = "response",
  show_ci = TRUE,
  xlab = "Age (years)",
  ylab = "Predicted total caloric intake (kcal)",
  legend_title = "Carbohydrate (percentiles)"
)

head(plot_data_list$plot_data)
```

## References
