---
title: "Formatting Multiple Imputation Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting Multiple Imputation Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
link-citations: yes
---

```{r setup}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to use the **svypooled()** function to format pooled regression model outputs from **mice** package objects into clean, fallacy-safe tables.

## Formatting Results from Multiple Imputation

Beyond survey data, a common analytical challenge is presenting results from models fitted to multiply imputed datasets. The **svyTable1** package provides the `svypooled()` function to address this. It takes a pooled model object (`mipo`) from **mice** and produces publication-ready regression tables.

A key feature of `svypooled()` is its ability to create a fallacy-safe table. This approach displays results for the primary exposure only while listing adjustment covariates in a footnote, preventing misinterpretation of covariate p-values [@westreich2013table].

## Example: Handling Missing Data in a Survey Design

The following sections illustrate the recommended workflow using the bundled
`nhanes_mortality` dataset (a processed NHANES cohort of diabetic adults).

### Data Preparation

```{r packages, warning = FALSE, message = FALSE}
library(svyTable1)
library(mice)
library(survey)
library(knitr)

options(survey.lonely.psu = "adjust")
```

We build an analytic dataset, define a binary hypertension outcome, and -- so
the imputation step has something to do -- induce some missingness in a
covariate.

```{r data}
data(nhanes_mortality)
nhanes_mortality$htn01 <- as.numeric(nhanes_mortality$htn == "Yes")

analytic <- nhanes_mortality[, c("htn01", "sex", "age", "smoking",
                                 "psu", "strata", "survey_weight")]

set.seed(123)
analytic$age[sample(nrow(analytic), 200)] <- NA
```

## Pooling from Imputed Data

We run `mice()`, fit a survey-weighted logistic regression model to each imputed dataset using `svyglm()`, and combine the results using `pool()` (Rubin's rules). This pooled model object is accepted directly by `svypooled()`.

A small number of imputations is used here only to keep the vignette fast; for a real analysis use `m >= 20` and check convergence.

```{r pool}
imputed_data <- mice(analytic, m = 2, maxit = 2, seed = 123, printFlag = FALSE)

fit_list <- list()
for (i in seq_len(imputed_data$m)) {
  completed_data <- complete(imputed_data, i)
  design_i <- svydesign(
    id = ~psu,
    strata = ~strata,
    weights = ~survey_weight,
    nest = TRUE,
    data = completed_data
  )
  fit_list[[i]] <- svyglm(
    htn01 ~ sex + age + smoking,
    design = design_i,
    family = quasibinomial()
  )
}

pooled_results <- pool(fit_list)
```

## Generating Tables with `svypooled()`

### Fallacy-Safe Table

This is the recommended format [@westreich2013table]. It presents results only for the main exposure and lists adjustment variables as a single footnote.

```{r t2}
svypooled(
 pooled_model = pooled_results,
 main_exposure = "sex",
 adj_var_names = c("age", "smoking"),
 measure = "OR",
 title = "Adjusted Odds of Hypertension (Fallacy-Safe)"
)
```

### Full Table (Appendix-Style)

This version displays results for every variable in the model. It is useful for supplemental materials or internal reporting but should not be used for primary interpretation.

```{r t2pool}
svypooled(
 pooled_model = pooled_results,
 main_exposure = "sex",
 adj_var_names = c("age", "smoking"),
 measure = "OR",
 title = "Adjusted Odds of Hypertension (Full Table for Appendix)",
 fallacy_safe = FALSE
)
```

## References
